The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-06 analysis/paper/paper.qmd

# Load packages ----
## List packages to be loaded
packages <- c(
  "here",                                       # Managing paths
  "tidyverse", "reshape2", "widyr", "readxl",  # Data manipulation and visualisation
  "furrr",                                      # Parallel use of purrr::map() functions
  "tidytext", "SnowballC", "tm",                # Text processing and analysis
  "topicmodels", "LDAvis",                      # Topic modeling
  "tidygeocoder",                               # Working with spatial data
  "igraph", "ggraph"                            # Network analysis and visualization
)

@@

# Add cleaned text column with removed stop words
emu$text_clean <- emu$text_clean |>
  str_remove_all(find_meta_stopwords(emu)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

@@

## Topic model

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.

@@

# Prepare document-term matrix as input for LDA model
emu_dtm <- convert_to_dtm(emu_words, "title", "word")

# Create LDA model
emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))

@@

```{r topic-model-beta}

# Extract topics as mixtures of terms from LDA model
emu_topics <- tidy(emu_lda, matrix = "beta")

# The top n most common words in each topic
emu_top_terms <- get_top_words_per_topic(emu_topics, 20)

@@

# TODO can this be generalised to more than two topics?
beta_spread <- emu_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

>>>>

# Load packages ----
## List packages to be loaded
packages <- c(
  "here",                                       # Managing paths
  "tidyverse", "reshape2", "widyr", "readxl",   # Data manipulation and visualisation
  "furrr",                                      # Parallel use of purrr::map() functions
  "tidytext", "SnowballC", "tm",                # Text processing and analysis
  "topicmodels", "LDAvis", "tsne",              # Topic modeling
  "tidygeocoder",                               # Working with spatial data
  "igraph", "ggraph"                            # Network analysis and visualization
)

@@

# Add cleaned text column with removed stop words
emu$text_clean <- emu$text_clean |>
  str_remove_all(find_meta_stopwords(emu)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))

@@

### Topic model

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.

The number of topics was qualitatively determined

```{r topic-model-lda}
#| dependson: short-stop-words

# Prepare document-term matrix as input for LDA model

# Reduce the size of the DTM by excluding words with a frequency lower than 5
emu_dtm <- convert_to_dtm(emu_theses_words, 
                          title_col = "title", 
                          word_col = "word", 
                          min_term_freq = 5)

# Fit LDA model with K clusters
# # Determine K value empirically
# K = get_number_of_topics(emu_dtm)
# # This returns K = 23
# TODO run the model with K = 23 and compare with results using qualitative value

# Determine K value qualitatively
# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
               control = list(seed = 2023, iter = 500, verbose = 25))

# Extract beta and theta statistics from LDA model
beta <- posterior(emu_lda)$terms
theta <- posterior(emu_lda)$topics

# Preview 10 terms from each topic
terms(emu_lda, 10)

# Add pseudo-names to topics based on the top 5 words in each topic
topic_names <- name_topics(beta)

@@

# What is the dominant topic in each document?
N = 20
topic_prop_examples <- theta[1:N,]
colnames(topic_prop_examples) <- topic_names
vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
                           document = factor(rownames(topic_prop_examples))), 
                     variable.name = "topic", id.vars = "document")  
ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ document, ncol = 5) +
  theme(legend.position = "none")

@@

```{r}

# What are the most probable topics in the entire collection?
topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
names(topicProportions) <- topic_names     # assign the topic names we created before
sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order

soP <- sort(topicProportions, decreasing = TRUE)
paste(round(soP, 5), ":", names(soP))

# How many documents are with the highest probability in each topic?
K = 5
countsOfPrimaryTopics <- rep(0, K)
names(countsOfPrimaryTopics) <- topic_names
for (i in 1:nDocs(emu_dtm)) {
  topicsPerDoc <- theta[i, ] # select topic distribution for document i
  # get first element position from ordered list
  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
}
sort(countsOfPrimaryTopics, decreasing = TRUE)

so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
paste(so, ":", names(so))
```

```{r}
topicToFilter <- 6  # you can set this manually ...
# ... or have it selected by a term in the topic name (e.g. 'children')
topicToFilter <- grep('river', topic_names)[1] 
topicThreshold <- 0.2
selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
filteredCorpus <- emu_theses[selectedDocumentIndexes,]
# show length of filtered corpus
filteredCorpus
```

```{r}

N = 82
topic_prop_per_year <- theta[1:N,]
colnames(topic_prop_per_year) <- topic_names

vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
                           document = factor(rownames(topic_prop_per_year))), 
                     variable.name = "topic", id.vars = "document") |> 
    left_join(select(emu_theses, title, "year" = graduation_year), by = c("document" = "title"))

# Plot topic proportions per year
vizDataFrameWithYear |> 
  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_viridis_d(name = "Topics") +
  xlab("Graduation year") +
  ylab("Proportion of topics") +
  labs(title = "Topic proportions per year")

```{r lda-vis}

# TODO check if this can be included in the Shiny dashboard
# Function to approximate the distance between topics
svd_tsne <- function(x) tsne(svd(x)$u)

# Convert DTM into JSON required by the LDAvis package
json <- LDAvis::createJSON(
  phi = beta, 
  theta = theta, 
  doc.length = rowSums(as.matrix(emu_dtm)), 
  vocab = colnames(emu_dtm), 
  term.frequency = colSums(as.matrix(emu_dtm)),
  mds.method = svd_tsne,
  plot.opts = list(xlab="", ylab="")
)

# Visualise topics model with LDAvis
LDAvis::serVis(json)

```{r topic-model-beta}
#| dependson: topic-model-lda

# Extract topics as mixtures of terms from LDA model
beta_tidy <- tidy(emu_lda, matrix = "beta")

# The top n most common words in each topic
emu_theses_top_terms <- get_top_words_per_topic(beta_tidy, 20)

@@

# TODO can this be generalised to more than two topics?

beta_spread <- emu_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))
